High-Dimensional Regularized Discriminant Analysis

### 2013-05-10 reports/paper-hdrda.Rnw
It helps to spell HDRDA right in the paper's filename.

For the microarray data sets, we computed error-rate estimates for the competing
classifiers after reducing the dimension of the data sets with feature selection
to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
\cite{Xu:2009fl}. To compare misclassification rates between different values of
$d$, we applied each value of $d$ considered to the same random allocation of
training and test data. Hence, for a given random partition, the set of features
selected for a small value of $d$ is a proper subset of the features selected
for a larger value of $d$. We report the average of the error-rate estimates for
each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
were no greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(), size = 1, linetype = 2)
p <- p + geom_point(aes(shape = Classifier), size = 6)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 20))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + theme(strip.text.x = element_text(size = 18))
p + theme(legend.position = "none")
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. The classifiers are marked as follows:
  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (square
  with X), and ours (asterisk). Approximate standard errors are no greater than
  $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
optimal estimates of $\lambda$ and $\gamma$ in the Supplementary Material. For
each of the four data sets, the distribution of the optimal values of $\lambda$
selected through cross-validation often had a notable bathtub shape, resembling
a Beta distribution with both shape parameters less than unity. Specifically,
the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
to quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{HDRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of our proposed classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious assumption of equal covariance matrices.

>>>>

For the microarray data sets, we computed error-rate estimates for the competing
classifiers after reducing the dimension of the data sets with feature selection
to $d = 100, 200, \ldots, 1000$, similar to the values of $d$ considered by
\cite{Xu:2009fl}. To compare misclassification rates between different values of
$d$, we applied each value of $d$ considered to the same random allocation of
training and test data. Hence, for a given random partition, the set of features
selected for a small value of $d$ is a proper subset of the features selected
for a larger value of $d$. We report the average of the error-rate estimates for
each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
were no greater than \Sexpr{max_se}.

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(), size = 1, linetype = 2)
p <- p + geom_point(aes(shape = Classifier), size = 6)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 20))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + theme(strip.text.x = element_text(size = 18))
p + theme(legend.position = "none")
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. The classifiers are marked as follows:
  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (square
  with X), and ours (asterisk). Approximate standard errors are no greater than
  \Sexpr{max_se}.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
optimal estimates of $\lambda$ and $\gamma$ in the Supplementary Material. For
each of the four data sets, the distribution of the optimal values of $\lambda$
selected through cross-validation often had a notable bathtub shape, resembling
a Beta distribution with both shape parameters less than unity. Specifically,
the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
to quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{HDRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of our proposed classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious assumption of equal covariance matrices.
